---
title: "daisugi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{daisugi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Getting Started with daisugi 

daisugi uses (x, y) nomenclature for fitting machines & primarily focus is for classification and regression tasks. to fit a machine, typically these machines are prefixed by `grow_` and to infer with a **grown** machine, one will `harvest_`. 

### A Classification Task

showcasing how to use daisugi for classification. our dataset comes from forested, a tabular data repo which lists forest attributes & whether an area is "forested" or "non-forested". 

```{r example}
library(daisugi)
library(forested)
library(rsample)

# defining splits for training and testing datasets
splits <- rsample::initial_split(forested::forested)
training <- rsample::training(splits)
testing <- rsample::testing(splits)


# (x, y) Training:
x_train <- training |>
  # target (and factors as not all engines handle cats)
  dplyr::select(-forested, -tree_no_tree, -land_type, -county)

# our target variable:
y_train <- training |> dplyr::select(forested) |> dplyr::pull()

# (x, y) Testing:
x_test <- testing |>
  dplyr::select(-forested, -tree_no_tree, -land_type, -county)

y_test <- testing |> dplyr::select(forested) |> dplyr::pull()

head(y_test)
```

#### yggdrasil decision forests  

YDF provides various methods such as boosters, decisions, forests. what is unique is the a oblique random split technique. daisugi provides the boosted method with yggdrasil.

```{r}
ydf_trees <- grow_yggdrasil_trees(
  x_train,
  y_train,
  trees = 20L
)

harvest_yggdrasil_trees(ydf_trees, x_test) |> head()
```

### snap boosting machines

SnapML is a IBM ML repository. within it, contains various ML engines, including their own implementation called SnapBoost. SnapBoost differs from XGBoost by using Heterogeneous Newton Boosting. In addition, for each iteration, SnapBoost method randomly chooses whether to use a decision tree with variable depth or a linear regressor with random fourier features. 

```{r}
snap_trees <- grow_snap_trees(
  x_train,
  y_train,
  trees = 10L
)

harvest_snap_trees(snap_trees, x_test) |> head()
```

### perpetual 

Perpetual is a budget-based boosting methodology. perpetual is designed to be a drop-in automl-like booster that does not require hyperparameter optimization. The idea being, a user can increase the budget, the 'predictive power' parameter, until loss plateaus. 

```{r}
perpetual_trees <- grow_perpetual_trees(
  x_train,
  y_train
)

harvest_perpetual_trees(perpetual_trees, x_test) |> head()
```

### wildwood

WildWood is a *new (2021)* & *advanced* random forest algorithm. ... "predictions produced by WildWood are an aggregation with exponential weights (computed on out-of-bag samples) of the predictions given by all the possible prunings of each tree." which differs from a standard random forest.

```{r}
wild_trees <- grow_wild_trees(
  x_train,
  y_train,
  trees = 10L
)

harvest_wild_trees(wild_trees, x_test) |> head()
```

### explainable boosting machines

"Explainable Boosting Machine (EBM) is a tree-based, cyclic gradient boosting Generalized Additive Model with automatic interaction detection." EBMs are considered 'glassbox' as they are easier to interpret, more effective than traditional GAMS, and "...often as accurate as SOTA blackbox models" while maintaining interpetability. 

```{r}
explainable_trees <- grow_explainable_trees(
  x_train,
  y_train,
  trees = 100L
)

harvest_explainable_trees(wild_trees, x_test) |> head()
```

### natural gradient boosting machines

NGBoost from Stanford ML Group. bringing uncertainty and probabilistic estimation to gradient boosting.  

```{r}
natural_trees <- grow_natural_trees(
  x_train,
  y_train,
  trees = 100L
)

harvest_natural_trees(natural_trees, x_test) |> head()
```
