---
title: "glossary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{glossary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(daisugi)
```

# Glossary 
daisugi is designed as an engine curation metarepo. there are a handful of nomenclature below that will be helpful for navigating the package. all machines are fitted with (x, y) interfaces. the machines are dictated with a prefix verb. Those are listed below: 

## Keywords
#### fit machines with: *grow_*
#### infer machines with: *harvest_*


## Machine Information

in addition to the typical reference page, this should be helpful for practitioners who would like to dive a bit deeper into the root repos, whitepapers, or other information. 

```{r}
#| include: false
# list source, alias, haikus ------------------

ebm <- c(
  "https://interpret.ml/docs/ebm.html",
  "explainable",
  "each feature speaks alone \n the sum is the whole model \n darkness becomes light"
)

perpetual <- c(
  "https://perpetual-ml.github.io/perpetual/",
  "perpetual",
  "one knob, the budget \n self-generalizing design \n no search, just descent"
)

snap <- c(
  "https://ibmsoe.github.io/snap-ml-doc/v1.6.0/manual.html#snapboost",
  "snap",
  "depth rolls like dice \n decision or linear \n random number tricks"
)

wild <- c(
  "https://github.com/pyensemble/wildwood",
  "wild",
  "every pruning weighed \n out-of-bag shadows voting \n wilder than the rest"
)

yggdrasil <- c(
  "https://github.com/google/yggdrasil-decision-forests",
  "yggdrasil",
  "sacred trunk rises \n c++ decision trees \n fast quiet giant"
)

# combine -------------------------------------------

df <- rbind(
  ebm = ebm,
  perpetual = perpetual,
  snap = snap,
  wild = wild,
  yggdrasil = yggdrasil
)

colnames(df) <- c("docsource", "alias", "haiku")

df <- as.data.frame(df, stringsAsFactors = FALSE)
```



```{r}
knitr::kable(df, format = "simple")
```